Exploring prosperLoanData by Tang Shuran
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.
library(base)
library(stats)
library(dplyr)
library(ggplot2)
library(knitr)
library(lubridate)
library(maps)
library(reshape2)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
setwd("D:/github/udacity_data_analyse/project4/report")
df<- read.csv("prosperLoanData.csv")
```

# Univariate Plots Section
```{r echo=FALSE, Univariate_Plots}
#interested feature: LoanStatus
ggplot(subset(df,!is.na(LoanStatus)),aes(x=LoanStatus)) + 
  geom_bar()
ggplot(subset(df,!is.na(LoanStatus)),aes(x=LoanStatus)) + 
  geom_bar(aes(y=..count../sum(..count..))) +
  ggtitle("Proportion of Loans by Status") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
summary(df$LoanStatus)

```



```{r}
#interested feature: EmploymentStatus
ggplot(subset(df,!is.na(EmploymentStatus)),aes(x=EmploymentStatus)) + 
  geom_bar(aes(y=..count../sum(..count..))) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
summary(df$EmploymentStatus)

```

```{r}
#interested feature: EmploymentStatus
ggplot(df,aes(x=EmploymentStatus)) + 
  geom_bar(aes(y=..count../sum(..count..))) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
summary(df$EmploymentStatus)
```

```{r}
table(df$Term)
table(df$IncomeRange)
summary(df$LoanOriginalAmount)
df$LoanOriginalAmount_cut<-cut(df$LoanOriginalAmount,c(0,1000,4000,6500,8500,12000,18000,25000,35000))
ggplot(df,aes(x=LoanOriginalAmount_cut)) + 
  geom_bar()

```

```{r}
qplot(x=log10(AvailableBankcardCredit+1),data=subset(df,!is.na(AvailableBankcardCredit)))+
  #scale_x_continuous(limits = c(0,200000),breaks=seq(0,200000,100))+
  facet_wrap(~IsBorrowerHomeowner)
```


```{r}
qplot(x=CurrentDelinquencies,data=subset(df,!is.na(CurrentDelinquencies)))+
  scale_x_continuous(limits = c(0,83),breaks=seq(0,83,5))

```


# Univariate Analysis
```{r}
str(df)

```


### What is the structure of your dataset?
LoanOriginalAmount: 
 Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
   1000    4000    6500    8337   12000   35000 
Term:
   12    36    60 
 1614 87778 24545 
AvailableBankcardCredit:
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max.    NA's 
      0     880    4100   11210   13180  646300    7544 
LoanStatus: Cancelled, Chargedoff,Completed,Current,Defaulted, FinalPaymentInProgress, Past Due
EmploymentStatus:Employed, Full-time, Not available, Not employed, Other, Part-time, Retired, Self-employed
ProsperRating..Alpha.: AA, A, B, C, D, E, HR


### What is/are the main feature(s) of interest in your dataset?
LoanOriginalAmount,IsBorrowerHomeowner,AvailableBankcardCredit,StatedMonthlyIncome,Occupation,EmploymentStatus

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?
LoanOriginationDate,Term,MonthlyLoanPayment,BorrowerAPR,LP_ServiceFees,ProsperRating

### Did you create any new variables from existing variables in the dataset?
Yes:
df$LoanOriginalAmount_cut<-cut(df$LoanOriginalAmount,c(0,1000,4000,6500,8500,12000,18000,25000,35000))

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?
I log-transformed the AvailableBankcardCredit. 
```{r}
qplot(x=log10(AvailableBankcardCredit+1),data=subset(df,!is.na(AvailableBankcardCredit)))

setwd("D:/github/udacity_data_analyse/project4/report")
df<- read.csv("prosperLoanData.csv")
df.Credit_and_income_by_Occupation_owner.long<-df%>%
  filter(!is.na(IsBorrowerHomeowner))%>%
  filter(!is.na(AvailableBankcardCredit))%>%
  filter(!is.na(Occupation))%>%
  group_by(Occupation,IsBorrowerHomeowner)%>%
  summarise(mean_Credit=mean(AvailableBankcardCredit),
            mean_income=mean(StatedMonthlyIncome),
            median_income=median(StatedMonthlyIncome),
            n=n())%>%
  ungroup()%>%
  arrange(Occupation)
df.meanCredit_by_Occupation_owner.wide<-dcast(
  df.Credit_and_income_by_Occupation_owner.long,
  Occupation~IsBorrowerHomeowner,
  value.var ="mean_Credit" )
df.meanIncome_by_Occupation_owner.wide<-dcast(
  df.Credit_and_income_by_Occupation_owner.long,
  Occupation~IsBorrowerHomeowner,
  value.var ="mean_income" )
df.medianIncome_by_Occupation_owner.wide<-dcast(
  df.Credit_and_income_by_Occupation_owner.long,
  Occupation~IsBorrowerHomeowner,
  value.var ="median_income" )


```



# Bivariate Plots Section
```{r echo=FALSE, Bivariate_Plots}
df.big_sample_Occupation<-df%>%
  filter(!is.na(Occupation))%>%
  group_by(Occupation)%>%
  summarise(n=n())%>%
  ungroup()%>%
  arrange(desc(n))
df.big_sample_Occupation[1:10,]
interested_occupations<-c("Computer Programmer","Teacher","Administrative Assistant","Analyst","Sales - Commission")
ggplot(aes(x=Occupation,y=StatedMonthlyIncome),data=subset(df,Occupation %in%interested_occupations))+
  geom_boxplot()+
  ylim(0,35000)
ggplot(aes(x=median_income,y=mean_Credit),data = df.Credit_and_income_by_Occupation_owner.long)+
  geom_point(alpha=0.5,color=I("red"),size=2)

df$ProsperRating..Alpha. <- factor(df$ProsperRating..Alpha.,levels = c('AA','A',"B","C","D","E","HR",""),ordered = TRUE)
ggplot(aes(x=ProsperRating..Alpha.,y=LoanOriginalAmount),data = subset(df,!is.na(df$ProsperRating..Alpha.)))+
  geom_boxplot()
ggplot(aes(x=ProsperRating..Alpha.,y=BorrowerAPR),data = subset(df,!is.na(df$ProsperRating..Alpha.)))+
  geom_boxplot()

```

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?
```{r}
cor.test(df.Credit_and_income_by_Occupation_owner.long$median_income,df.Credit_and_income_by_Occupation_owner.long$mean_Credit)

```
Generally speaking, the loan original Amount tends to be larger as the ProsperRating goes higher.
The loan original amount squeezes a lot when the prosper rating drop from C to D.
The gap between other ratings are not that huge.

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?
The loanee with home ownership generally has a higher credit and state income than the ones without ownership.
The state monthly income of loanees vary a lot among different occupations. Even within the same occupation, the there is huge variance for some particular occupation. Some outlier can reach even 3 or 4 times as the mean income.
### What was the strongest relationship you found?
For a certain occupation, the mean credit and median come are strongly related.
The correlation between them is as high as 0.799.
The loanee has to pay more interests if the loan has a lower ProsperRating.


# Multivariate Plots Section


```{r echo=FALSE, Multivariate_Plots}
ggplot(aes(x=BorrowerAPR,y=LP_InterestandFees),data=subset(df,!is.na(ProsperRating..Alpha.)))+
  geom_point(alpha=0.4,aes(color=ProsperRating..Alpha.),position = "jitter")

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection
